---
title: "Assignment-3"
author: "V Rana Prudhvi Tej"
date: "2023-11-05"
output:
  html_document: default
  pdf_document: default
---

```{r}
## Load required libraries
library(tidyverse)
library(reshape)
library(reshape2)
library(caret)
library(e1071)

## Load the data csv file
Uni_Bank <- read.csv("/Users/ranatej/Downloads/UniversalBank-1.csv")

head(Uni_Bank)
tail(Uni_Bank)
```

```{r}
## defining Column names
colnames(Uni_Bank)

#Data transformation into factors and categories.

Uni_Bank$`Personal Loan` = as.factor(Uni_Bank$Personal.Loan)
Uni_Bank$Online = as.factor(Uni_Bank$Online)
Uni_Bank$CreditCard = as.factor(Uni_Bank$CreditCard)
```

```{r}
#Dividing the data in half, with 40% going into the validation set and 60% going into the training set
set.seed(456)

# 60 % training data and 40% Validation data respectively
Uni_Bank.train.data <- sample(row.names(Uni_Bank), 0.6*dim(Uni_Bank)[1])
Uni_Bank.valid.data <- setdiff(row.names(Uni_Bank), Uni_Bank.train.data)

#putting training and validation data into separate dataframes from the dataset
Uni_Bank.train <- Uni_Bank[Uni_Bank.train.data, ]
Uni_Bank.valid <- Uni_Bank[Uni_Bank.valid.data, ]

# Duplicating the data frame UB.train
train <- Uni_Bank[Uni_Bank.train.data, ]
# Duplicating the data frame UB.valid
valid <- Uni_Bank[Uni_Bank.train.data,]

```

## A. Create a pivot table for the training data with Online as a column variable, CC as a row variable, and Loan as a secondary row variable. The values inside the table should convey the count. In R use functions melt() and cast(), or function table().

```{r}
## melt data from data
m_data = melt(train,id=c("CreditCard","Personal.Loan"),variable= "Online")

# Casting the melted data
c_data <- dcast(m_data, CreditCard + Personal.Loan ~ value, fun.aggregate = length)

c_data[,c(1,2,3,14)]
```

#B. Consider the task of classifying a customer who owns a bank credit card and is actively using online banking services. Looking at the pivot table, what is the probability that this customer will accept the loan offer? [This is the probability of loan acceptance (Loan = 1) conditional on having a bank credit card (CC = 1) and being an active user of online banking services (Online = 1)].

```{r}
#The pivot table indicates that the value for CC is and the value for Loan is 89. 
Uni_Bank.Loan.CC1 <- 89/3000

# which is 29 %.
Uni_Bank.Loan.CC1
```
## C. Create two separate pivot tables for the training data. One will have Loan (rows) as a function of Online (columns) and the other will have Loan (rows) as a function of CC.

```{r}
## transforming the train data frame into a lengthy format, using "Online" as a variable to be melted and "Personal.Loan" as an identification
m1 = melt(train,id=c("Personal.Loan"),variable = "Online")

 ## creating a long format out of the train data frame, using "CreditCard" as an identifier and "Online" as a variable that needs to be melted
m2 = melt(train,id=c("CreditCard"),variable = "Online")
```

```{r}
# Casting Personal loan and online values
c1=dcast(m1,`Personal.Loan`~Online) 

# Casting Personal loan and online values
c2=dcast(m2,CreditCard~Online)
```

```{r}
#displays the quantity of personal loans compared to online
Uni_Bank.Loanonline=c1[,c(1,13)]
Uni_Bank.LoanCC = c2[,c(1,14)]
Uni_Bank.Loanonline
```
```{r}

# The quantity of credit cards in relation to online is indicated by #.
Uni_Bank.LoanCC

```
## D. Compute the following quantities [P (A | B) means “the probability of A given B”]:
#1.P (CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors) #2.P(Online=1|Loan=1) 
#3.P (Loan = 1) (the proportion of loan acceptors)
#4.P(CC=1|Loan=0)
#5.P(Online=1|Loan=0)
#6.P(Loan=0)

```{r}
# making a pivot table where personal loans are represented by columns 14 and 10.
table(train[,c(14,10)])
```
```{r}
# Making a pivot table for the online and personal loan columns 13 and 10.
table(train[,c(13,10)])
```

```{r}
# Pivot table for personal loans There are, from training, 2725 and 275, respectively.
table(train[,c(10)])
```
```{r}
## By consulting the above p, we may determine the CC= 1 and Loan = 1 values.
Uni_Bank.CCUni_Bank.Loan1 = 89/(89+200)
Uni_Bank.CCUni_Bank.Loan1
```
2. P(Online=1|Loan=1)

```{r}
##The pivot table above UB.ONUB.Loan1 gives us the data for online = 1 and loan = 1.
Uni_Bank.ONUni_Bank.Loan1 =177/(177+112)
```


3. P(Loan = 1)

```{r}
 # By referring the above pivot table we can get the Loan = 1
Uni_Bank.Loan1 =289/(289+2711)
Uni_Bank.Loan1
```


4. P(CC=1|Loan=0)

```{r}
#The CC = 1 and Loan = 0 values can be obtained by using the pivot table above.
Uni_Bank.CCLoan.01= 794/(794+1917) 
Uni_Bank.CCLoan.01
```

5. P(Online=1|Loan=0)

```{r}
# The pivot table above gives us the numbers for online = 1 and loan = 0.
Uni_Bank.ON1.L0= 1665/(1665+1046)
Uni_Bank.ON1.L0
```

6. P(Loan=0)

```{r}
# The pivot table above allows us to extract the Loan = 0 values.
Uni_Bank.Loan0= 2711/(2711+289)
Uni_Bank.Loan0
```
##E. Use the quantities computed above to compute the naive Ba1 probability P(Loan = 1 | CC = 1, Online = 1).

```{r}
# Given probabilities
P_CC_Loan1 <- 0.096
P_Online_Loan1 <- 0.833
P_L1 <- 0.0125

# Calculate Naive Bayes probability P(Loan = 1 | CC = 1, Online = 1)

Uni_Bank_Naivebayes <- (P_CC_Loan1)*(P_Online_Loan1)*(P_L1)
Uni_Bank_Naivebayes

```

## F. Compare this value with the one obtained from the pivot table in (b). Which is a more accurate estimate?

```{r}
# Naive Bayes Probability (from calculation in E)
naive_bayes_prob <- 0.0009996

# Pivot Table Probability
pivot_table_prob <- 0.02966667

# Compare the odds and print a message stating which is more likely to be true.
if (naive_bayes_prob > pivot_table_prob) {
  message("Naive Bayes Prob is more accurate: ", naive_bayes_prob)
} else if (naive_bayes_prob < pivot_table_prob) {
  message("Pivot Table Prob is more accurate: ", pivot_table_prob)
} else {
  message("Both Prob are the same: ", naive_bayes_prob)
}

```
##Based on the comparison, the pivot table probability (0.02966667) is considered more accurate compared to the Naive Bayes probability


## G. Which of the entries in this table are needed for computing P(Loan = 1 | CC = 1, Online = 1)? Run naive Bayes on the data. Examine the model output on training data, and find the entry that corresponds to P(Loan = 1 | CC = 1, Online = 1). Compare this to the number you obtained in (E).

```{r}

names(Uni_Bank)

names(Uni_Bank.train)

# Choose the pertinent training columns.
Uni_Bank.train <- Uni_Bank[, c("CreditCard", "Online", "Personal Loan")]

#Change the columns' names to eliminate gaps
colnames(Uni_Bank.train) <- c("CreditCard", "Online", "PersonalLoan")

# Convert "Online" and "CreditCard" to factors with the proper levels
Uni_Bank.train$CreditCard <- factor(Uni_Bank.train$CreditCard, levels = c(0, 1), labels = c("No", "Yes"))
Uni_Bank.train$Online <- factor(Uni_Bank.train$Online, levels = c(0, 1), labels = c("No", "Yes"))
```




```{r}
# Print the probability
print("Prob of Loan = 1 given CC = 1 and Online = 1:")
```



