---
title: "FML ASSIGNMENT 4"
author: "V RANA PRUDHVI TEJ"
date: "2023-11-19"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

## SUMMARY 

For this project, I'll use the k-implies grouping approach to conduct a non-various levelled bunch inspection. To enable the extraction of crucial information, the data will be divided into consistent categories. The first dataset and the necessary bundles should be stacked first. Data from roughly twenty-one pharmaceutical companies are included.

The rationale behind the choices of market capitalization, leverage, rev growth, net profit margin, beta, PE ratio, ROE, and ROA
The variables that were selected are common financial measures that are employed in business performance evaluation and comparison. Market Cap, Beta, PE Ratio, ROE, ROA, Turnover of Assets, Leverage, Rev Growth, and Net Profit Margin are a few of them. When taken as a whole, these variables provide a complete view of a company's productivity, profitability, and stability.


1. The market capitalization of pharmaceutical firms ranges from 0.41 to 199.47, indicating their overall size and worth.


2. Beta: runs from 0.18 to 1.11 and shows how responsive a company's returns are to shifts in the market.


3. PE Ratio: reflects the stock price of a corporation in relation to its earnings. It can fall between 3.6 to 82.5.


4. ROI  : ranges from 3.9 to 62.9, indicating how well a business uses shareholder equity to generate profits.


5. ROA  : 0.3 to 1.1. Assesses an organization's ability to generate revenue from its resources.


6. Asset Turnover: shows how well a business generates income from its assets. The range is 0.5 to 1.1.


7. Leverage: demonstrates the amount of debt a company employs to fund its activities; spans the interval 0–3.51.


8. Rev_Growth: varies from -3.17 to 34.21 and displays the percentage change in revenue over a certain time period.


9. Net Profit Margin: This statistic, which spans from 2.6 to 25.54, indicates the proportion of revenue that is turned into profit.

Normalising the data:
The numerical variables need to be normalised in order for each variable to contribute proportionately to the clustering procedure. Since these variables may have multiple units or scales, normalisation helps prevent one variable from dominating the clustering based only on its magnitude. In comparison, Market Cap is in the hundreds, while Beta is a fraction between 0 and 1.

I've selected K-means over DBSCAN because it's commonly used in exploratory data analysis to identify patterns and groups within the data, and because K-means clustering can provide details about the financial profiles of pharmaceutical businesses. DBSCAN can detect groupings of enterprises with similar financial characteristics, which supports investment analysis and strategic decision-making. This makes it effective for datasets with dense areas. It is also easy to understand.The K-means algorithm requires a fixed number of clusters, k. The user can decide how many clusters to construct, which may be useful in some situations. The number of clusters may not be easily determined using DBSCAN or hierarchical clustering.

The dataset is divided into five groups according to numerical variables.
When interpreting each cluster, financial ratios and performance measurements are taken into account.
Cluster characteristics include things like net profit margin, sales growth, leverage, beta, ROA, and ROE.

cluster 1: From greatest to lowest, the hold, moderate buy, moderate sell, and moderate sell are arranged. They came from the US, the UK, and Switzerland, with the US having the highest ranking. They are listed on the NYSE.


Cluster 2: The firms are evenly distributed throughout AMEX, NASDAQ, and NYSE; the hold and moderate buy medians are different, with the hold being bigger than the moderate buy. The count from the US and Germany is also different.


cluster 3 is equally distributed throughout the US and Canada, has equal Hold and Moderate Buy medians, and is solely listed on the NYSE.


Cluster 4: This group is dispersed across the US and the UK, shares the same hold and moderate buy medians, and is listed on the NYSE.

Cluster 5 consists of equal parts moderate buy and moderate sell. It is listed on the New York Stock Exchange and spread throughout France, Ireland, and the United States.

We look at the relationships that exist between clusters and variables 10 to 12.
Bar plots are used to display the frequency distribution of non-clustered variables inside each cluster.Below the graph, using the bar graph, are the appropriate labels and an explanation.

## PROBLEM STATEMENT 

An equities analyst is studying the pharmaceutical industry and would like your help in exploring and understanding the financial data collected by her firm. Her main objective is to understand the structure of the pharmaceutical industry using some basic financial measures. Financial data gathered on 21 firms in the pharmaceutical industry are available in the file Pharmaceuticals.csv Download Pharmaceuticals.csv. For each firm, the following variables are recorded:  

1.Market capitalization (in billions of dollars)


2.Beta


3.Price/earnings ratio


4.Return on equity


5.Return on assets


6.Asset turnover


7.Leverage


8.Estimated revenue growth


9.Net profit margin


10.Median recommendation (across major brokerages)


11.Location of firm’s headquarters


12.Stock exchange on which the firm is listed

Use cluster analysis to explore and analyze the given dataset as follows: 

Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on. 
Interpret the clusters with respect to the numerical variables used in forming the clusters. Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used in forming the clusters)
Provide an appropriate name for each cluster using any or all of the variables in the dataset.





#installing the libraries using install.packages() and calling the requried libraries 
```{r}
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
library(ggplot2)
library(ISLR)
library(gridExtra)
library(flexclust)
library(dbscan)

```

```{r}
#importing the dataset and reading the dataset  
dataset <- read.csv("/Users/ranatej/Downloads/Pharmaceuticals.csv")
head(dataset)

```

##  1. To cluster the 21 enterprises, use just the number variables (1 to 9). Explain the decisions you made when performing the cluster analysis, including the weights assigned to the individual variables, the particular clustering algorithm(s) you chose, the number of clusters you produced, and so on.


```{r}
#To exclude any potential missing values from the data
Pdata <- na.omit(dataset)
#Assembling numerical data to group 21 businesses from columns 1 through 9.
row.names(Pdata)<- Pdata[,1]
Ph<- Pdata[, 3:11]
head(Ph)

```


```{r}
#normalizing the data using Scale function
ph2<- scale(Ph)
head(ph2)

```

```{r}
#K-means clustering computation in R for various centres
#Using several K values, compare the variations in the outcomes.

k <- kmeans(ph2, centers = 2, nstart = 30)
k1<- kmeans(ph2, centers = 5, nstart = 30)
k2<- kmeans(ph2, centers = 6, nstart = 30)
p1<-fviz_cluster(k, data = ph2)+ggtitle("k=2")
p2<-fviz_cluster(k1, data = ph2)+ggtitle("k=5")
p3<-fviz_cluster(k2, data = ph2)+ggtitle("k=6")
grid.arrange(p1,p2,p3, nrow = 2)


```


```{r}
#To determine the optimal radius or eps value.

#Graph to determine the optimal radius value at the minimum points of 4.
dbscan::kNNdistplot(Ph, k=4)
# DBSCAN Algorithm with minpts = 4 and eps = 30
dbscan_<- dbscan::dbscan(Ph, eps = 30, minPts = 4)

# Output from the groups
print(dbscan_)

# To determine which point is a member of which cluster
print(dbscan_$cluster)

# Visualization of clusters
fviz_cluster(dbscan_, Ph) + ggtitle("DBSCAN Plot")

```
#K-means clustering can provide information on the financial profiles of pharmaceutical companies, and it's widely used in exploratory data analysis to uncover patterns and groups in the data, which is why I choose it over DBSCAN. DBSCAN can be helpful in exposing groups of enterprises with similar financial features, which can aid in strategic decision-making and investment analysis for datasets with dense areas. It is also easy to understand.The K-means algorithm requires a fixed number of clusters, k. The user can decide how many clusters to construct, which may be useful in some situations. The number of clusters may not be easily determined using DBSCAN or hierarchical clustering.



```{r}
# Hierarchical Clustering

# Determine the data's euclidean distance.
euclidean_dis <- dist(ph2, method = "euclidean")

# Hierarchical Clustering
h <- hclust(euclidean_dis, method = "complete")

# Show the Dendrogram output at height=5.
plot(h, cex = 0.75, main = "Dendrogram of Hierarchical Clustering")
rect.hclust(h, h=5, border = 1:5)



```



```{r}
#Choosing the best clusters with the Elbow technique
distance <-  dist(ph2, method = "euclidean")
#distance matrix a data matrix's rows apart.
fviz_dist(distance)

```
#The total within-cluster sum of squares (wss) for each k should be determined. Tot.withinss is the total within-cluster sum of squares. Determine and graph the wss for k = 1 to k = 10 and extract the wss for clusters 2–15. It is commonly accepted that the position of a bend, or knee, in the plot indicates that k = 5 is the right number of clusters.


```{r}
set.seed(123)
wss<- function(k){
kmeans(ph2, k, nstart =10)$tot.withinss
}
k_values<- 1:10
wss_c<- map_dbl(k_values, wss)
plot(k_values, wss_c,
     type="b", pch = 16, frame = TRUE,
     xlab="Number of clusters",
     ylab="Total within-clusters sum of squares")

```
#The elbow at position 2 is visible in the graph above, but because the graphical representation is less acute, it is still confusing.




```{r}
#Using the following Silhouette technique
fviz_nbclust(ph2,kmeans,method="silhouette")

```

#The Silhouette approach will be employed due to its lucid depiction of K=5.
#Complete analysis, results extraction utilising five clusters, and results visualisation

```{r}
set.seed(123)
f_l<- kmeans(ph2, 5, nstart = 25)
print(f_l)
fviz_cluster(f_l, data = ph2)

```
##  2.Interpret the clusters with respect to the numerical variables used in forming the clusters. Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used in forming the clusters)

## Interpret the clusters with respect to the numerical variables used in forming the clusters

```{r}
# Interpreting the clusters in light of the numerical variables that were utilised to create them
Ph%>%
  mutate(Cluster = f_l$cluster) %>%
  group_by(Cluster)%>% summarise_all("mean")
  clusplot(ph2,f_l$cluster, color = TRUE, labels = 2,lines = 0)

```
Cluster 1: AHM, SGP, WYE, BMY, AZN, ABT, NVS, LLY - This group has the lowest sales growth and the highest net profit margin. These businesses have slow revenue growth and comparatively little leverage. businesses have the best return on equity and net profit margin, indicating that businesses offer a wide range of profitable products. Thus, they don't have to use up all of their resources. These businesses have little leverage because they don't need to borrow money from the capital market.
 
Cluster 2: BAY, CHTT, IVX - This cluster has poor ROA, revenue growth, and net profit margin, but high leverage and beta. These businesses are examples of creative startups in the sector. Compared to those well-known brands, they are relatively minor in terms of market capitalization, and few people are familiar with their name. Their revenue growth is modest and their net profit margins are low due to their recent establishment, lack of experience, and lack of profitable products that can produce cash flow. They have a low ROA and a high degree of leverage since they heavily rely on R&D. Since they have a high beta and are investing in the future, their price will increase in a rising market.

Cluster 3: PHA and AGN There are just two businesses in this cluster: AGN and PHA. Its net profit margin, low ROA, lowest beta, and highest P/E ratio are all present.Consequently, many businesses had modest net profits in the past but great expectations for the future. The market values them highly since they may invest a significant amount of money in D&I in cutting-edge technologies. Nevertheless, investors bear greater risk due to its high price.

The group with the largest market capitalization, high ROE and ROA, high net profit margin, high asset turnover, and low leverage is Cluster 4, which includes JNJ, MRK, PFE, and GSK. These businesses represent the industry leaders with their highest market capitalizations and most noticeable positions. These businesses have the lowest leverage values, high ROE, ROA, asset turnover, and outstanding capital usage. With every dollar invested in these companies, they stand to benefit the most. In addition to mature products that require little capital or asset investment from the companies but generate substantial revenue and strong net profit margins, they must have a few best-selling and dominant items in the market. Pfizer is one example of such a product.

Cluster 5: AVE, MRX, WPI, and ELAN  This cluster features low market capitalization, low P/E, low turnover rate, high beta, and strong revenue growth. These conventional small-sized enterprises appear to lack significant capital utilisation abilities based on their poor ROE, ROA, and turnover rates. Nevertheless, we can presume that the organisations are being led in the right way by either internal reformation or external market developments given the robust rate of revenue growth. Furthermore, the lowest P/E suggests that their share price is still modest.


## Do the clusters exhibit any patterns in relation to the numerical variables (10 to 12)? (those omitted from the cluster formation)

```{r}
p_clue <- dataset[12:14] %>% mutate(Clusters=f_l$cluster)
ggplot(p_clue, mapping = aes(factor(Clusters), fill =Median_Recommendation))+geom_bar(position='dodge')+labs(x ='Clusters',y ='Frequence')
ggplot(p_clue, mapping = aes(factor(Clusters),fill = Location))+
  geom_bar(position = 'dodge')+labs(x ='Clusters',y = 'Frequence')
ggplot(p_clue, mapping = aes(factor(Clusters),fill = Exchange))+geom_bar(position = 'dodge')+
  labs(x ='Clusters',y = 'Frequence')


```



## Cluster 1:


The median recommendation is that Cluster 1 has a solid hold.


Location: There are three locations in Cluster 1, more in the United States than in the United Kingdom or Switzerland.


Exchange: The NYSE, which has a sizable number of participants, is the only exchange in Cluster 1.



## Cluster 2:

Cluster 2 has a poor buy rating and a strong hold rating, according to the median recommendation.

Location: The US is ranked higher than Germany in two places within Cluster 2.

Exchange: There are three equally distributed exchanges in Cluster 2: AMEX, NASDAQ, and NYSE.


## Cluster 3:

Median suggestion: Based on the median suggestion, Cluster 3 has low purchase and low hold values.

Location: The United States and Canada are the only two equally distributed locations in Cluster 3.

Exchange: The NYSE is the only exchange in Cluster 3.


## Cluster 4:
Median advice: The median advice is that Cluster 4 has a high purchase and a high hold.

Location: There are two locations for Cluster 4, with the US having a significant advantage over the UK.

Exchange: The NYSE is the only exchange in Cluster 4.


## Cluster 5:

Cluster 5 has a moderate buy and moderate sell rating, which is the median suggestion.

Location: There are three locations for Cluster 5, the most notable being the United States.

Exchange: The NYSE is the only exchange in Cluster 5.






## 3.Provide an appropriate name for each cluster using any or all of the variables in the dataset.

 APPROPRIATE NAME :
 
 cluster 1 - Elevated hold cluster
 
 
 cluster 2 - Hold cluster
 
 
 cluster 3 - Cheapest cluster
 
 
 cluster 4 - Purchase hold  cluster
 
 
 cluster 5 - Purchase sell  cluster












































































































